Compositional Model-Checking for Fault-Tolerant Distributed System Implementations

### 2016-11-24 doc/om-paper/paper.tex
mucked with intro.

\begin{abstract}
With proof techniques like IC3 and $k$-induction with user-supplied invariants, model-checking scales further than ever before. Still, fault-tolerant distributed systems are particularly challenging to model-check given their large state space and non-determinism. The typical approach to controlling complexity is to construct ad-hoc abstractions of faults, message-passing, and behavior to simplify their behavior. However, these abstractions come at the price of divorcing the model from its implementation and making refactoring difficult. In this work, we present a systematic model for fault-tolerant distributed system verification that combines calendar automata, synchronous observers, and abstract state machines specialized for model-checking. We demonstrate the efficacy of the framework by presenting a the first model-checked implementation of the Oral Messages algorithm then demonstrate how modifications along axes such as the fault model, topology, node behavior, and require only local modifications to the model and proofs.
\end{abstract}

@@

\lee{needs to be redone.}
Fault-tolerant distributed systems are famously complex to design and build, yet are the  backbone of life-critical systems. Consequently, this class of systems demand high-assurance of correct design and implementation. Formal verification can be used to prove the correctness of their designs.

There is substantial literature on verifying fault-tolerant distributed systems. Most of this work uses interactive theorem-proving~\cite{}. Interactive theorem-proving is useful in the domain because of the complexity of the models and because distributed system verification is typically parameterized by the number of nodes; automated methods perform poorly along these two metrics. However, two problems with interactive verification are that it requires specialized expertise and it is labor intensive.

In contrast, model-checking for this class of systems is automated~\cite{}, but has limited scalability, particularly in the context of modeling the nondetermism of faults and distributed computation. The way that scalability is usually addressed is by specialization and ad-hoc model-specific abstractions. Specialization includes fixing a (small number) of nodes and topology for distributed protocols that are parameterized by the number of nodes and network.

Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value. To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$$ However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values). (2) Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided. While these optimizations might be safe under particular assumptions, they require careful meta-level reasoning and justification and divorce the model from any realization of it.

However, recent advances in model-checking allow for better scalability and model fidelity, and they begin to blur the lines between model-checking and interactive theorem-proving. In particular, $k$-induction for infinite-state systems allows the specifier to assert $k$-inductive lemmas

  (We discuss the applicability of property-directed reachability in Section~\ref{sec:related}.)

Our contributions combine a number of ideas in the formal verification literature to build scalable, generic formal models of fault-tolerant distributed systems in which verification is more automated than in traditional interactive theorem-proving. We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. The model combines several formalisms from the literature to efficiently but systematically model and verify a broad class of systems. The model extends \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal}. A calendar automata is a model inspired by discrete-event simulation, used in testing real-time systems~\cite{}.  Calendar automata have previously been used to verify bounded asynchronous systems~\cite{}; here we verify a synchronous system.

The model also employs \emph{synchronous observers}~\cite{}. Synchronous observers are a modeling abstraction in which an ``observer'' actor is synchronously composed with the system under verification. We use synchronous observers to compose distinct transition systems modeling the system, environment, assumptions, and requirements. Finally, we use an \emph{abstract state machine}~\cite{}, also composed with the system under verification using synchronous observation. The abstract state machine simplifies discovering invariants and debugging counterexamples.

We validate the model with a number of small examples \lee{todo}, as well as an extended case-study, verifying the Oral Messages (1) protocol (OM(1))~\cite{}. While our model is more detailed---it explicitly models channels---and also more general---it allows bounded asynchrony between processes---it is the most scalable model-checking verification of the protocol.

@@

>>>>

\begin{abstract}
With proof techniques like IC3 and $k$-induction with user-supplied invariants, model-checking scales further than ever before. Still, fault-tolerant distributed systems are particularly challenging to model-check given their large state space and non-determinism. The typical approach to controlling complexity is to construct ad-hoc abstractions of faults, message-passing, and behavior to simplify their behavior. However, these abstractions come at the price of divorcing the model from its implementation and making refactoring difficult. In this work, we present a systematic model for fault-tolerant distributed system verification that combines calendar automata, synchronous observers, and abstract state machines specialized for model-checking. We demonstrate the efficacy of the framework by presenting the first model-checked \emph{implementation} of the Hybrid Oral Messages algorithm, using a fast majority vote, then demonstrate how modifications along axes such as the fault model, topology, node behavior, and require only local modifications to the model and proofs.
\end{abstract}

@@

\lee{talk about Amazon/TLA? Paxos made live?}

Fault-tolerant distributed systems are famously complex, yet are the backbone of life-critical systems, such as commercial avionics. Consequently, this class of systems demands high-assurance of correct design and implementation. Formal verification can help provide that assurance.

The verification of this class of systems has usually been at the algorithmic level, eliding details about a concrete implementation, and historically, has relied on formal models verified by interactive theorem-proving~\cite{om-acl2-impl,Young97:IC,csl-93-2,pvs}. If formal verification is to be introduced into the workflow of system designers, though, we need automated methods and programmatic verification of \emph{implementations}. (Mostly) automated proof techniques are required to reduce the need for specialized verification expertise. System designers create software and hardware implementations to test, simulate, and deploy. Discrepancies between implementations can formal models can easily arise if the latter is abstracted too much from the former, particularly if those abstractions are ad-hoc and system specific. Furthermore, as an implementations are tweaked and modified to explore the design space, it is easy for the formal models to fall out of synch.

To address this problem, we envision building a domain-specific language (DSL) for specifying and exploring distributed fault-tolerant system implementations. We envision a DSL that generates both code for simulation, testing, and deployment, as well as formal models for verification (we say more about the DSL in Section~\ref{sec:conclusions}).

The first step in realizing this vision, however, and the one we address in this paper, is to develop an intermediate representation (IR) for the DSL to target when generating formal models. Our target for the IR is one supported by infinite-state model-checking. With proof techniques like IC3~\cite{} and $k$-induction~\cite{}, model-checking scales further than ever before. We envision an IR that allows for specifying system behavior, assumptions, and requirements, but also includes invariants that can be systematically generated by the DSL. Finally, the IR must support easy refactoring and modifications to the design. Making small changes to system specification---for example, modifying the fault model---should require only localized changes to the IR representation, including proofs. In short, the IR should be compositional.

In this work, we combine a number of ideas known in the literature and folklore to build scalable and compositional formal models of fault-tolerant distributed systems. We describe a generic formal model for fault-tolerant distributed systems that aims to reduce the need for ad-hoc verification abstractions and optimizations while maintaining scalability in Section~\ref{sec:model}. There are three aspects of the model: (1) \emph{calendar automata}, originally developed by Dutertre and Sorea~\cite{cal} for specifying real-time systems; (2) \emph{synchronous observers}, originally from the synchronous languages community~\cite{} and more recently popularized by Rushby for model-checking; and (3) \emph{abstract state machines} for abstracting complex systems.

To demonstrate the model, we describe a formally-verified implementation of the Oral Messages Hybrid algorithm (OMH)~\cite{}, implementing the Boyer-Moore Fast Majority Vote algorithm~\cite{}. The verification is interesting in its own right, as it is the first fully parametric (on the number of nodes) model-checked implementation of the classic algorithm. Then, we modify the algorithm along axes such as the fault model, topology, node behavior, and show that only local modifications are required to the model and proofs.

%% Two examples of model-specific abstractions include the following: (1) Suppose we model a system in which a process that broadcasts the values $\{0, 1\}$ in the nominal case, but when the process suffers a fault, it broadcasts no value. To model the system, it is typical to define an enumerated type of broadcast values as $$\{0, 1, NONE\}$$ However, the enumeration conflates the environment (i.e., faults) and system (i.e., the non-faulty broadcast values). (2) Another example is to model message passing as shared state between processes to reduce the number of state variables in the model, since the message passing mechanism is elided. While these optimizations might be safe under particular assumptions, they require careful meta-level reasoning and justification and divorce the model from any realization of it.

@@

\lee{talk about atom-sally}

\section*{Acknowledgments}
\lee{NASA contract, honeywell ppl, rushby's model?}
